The Analogue Method for Precipitation Forecasting: Finding Better Analogue Situations at a Sub-Daily Time Step

### 2015-12-17 Moving temporal window - HESS/template.tex
Removing the we...

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent to a finer resolution. Therefore, the analogues situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise we would not know what precipitation values to assign to them. However, we can expect that the analogy of the synoptic situations do not occur systematically at the same time of the day, and that better candidates can be found at different times. On this assumption, we introduced a sliding time window (STW) to allow searching for candidates at different hours of the day.

@@

One can question the interest of using daily totals when a 6-hourly precipitation series can be used. The first reason is that the 6-hourly time series generated by the AM may not represent well the dynamic of the accurate precipitation (results not shown), due to a smoothed signal. Then, sometimes we just don't need finer resolution than the daily time step. Finally, when using a reconstructed precipitation archive, the errors in intra-day precipitation distributions have a lesser impact.

@@

It can be assumed that the atmospheric conditions with a low dynamism, such as the frequent anticyclonic situations, will not be radically improved by the introduction of the STW. Conversely, dynamic situations, such as weather disturbances, have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure area and the fronts change significantly during a day. We can therefore expect to improve particularly these situations with a higher dynamism when introducing a STW, as we may find better matches to the target situation.

The dynamism of a given atmospheric condition cannot be easily quantified. We thus consider here the basic assumption that the more a day is rainy, the more dynamic the situation is. The results of this analysis are summarized in Figure \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation between two thresholds. The number of analogues being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations and confirms our intuition.

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the candidates situations are not more numerous, as we subsample in the previously $N_{1}$ selected analogues, but the dates may have changed. In contrast to earlier, both a reduction and an increase of the RMSE criterion values are possible. Figure \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

@@

A systematic improvement of the S1 values was previously observed. However, finding better analogue situations does not obligatory imply better forecast skills. Therefore, we have to assess the impact of the introduction of the STW, and thus the selection of other analogue dates, on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see section \ref{sec:data}). The target dates remain unchanged, since we keep the original time slot (6~h~UTC - 6~h~UTC the next day), and thus the performance scores can be directly compared with the former ones.

@@

The values of the CRPSS scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration. With the introduction of the STW, we get back the performance loss related to the reduction of the archive length. In our case, this improvement is approximately a doubling of the archive size.

@@

As we saw in the previous section, the performance improvement brought by the STW is lost when using a simplistic precipitation time series. In order to get closer to the chronology of the accurate precipitation, the reconstruction procedure must rely on some informative variable. Data from the NCEP/NCAR reanalysis 1 (section \ref{sec:data}) was used in this perspective, despite their rough resolution.

The first step consisted in determining which variables from the precipitable water or the relative humidity (at 1000~hPa, 925~hPa or 850~hPa) produces the precipitation chronology that is the most correlated (on non-zero values) with the accurate 6-hourly precipitation time series on the period 1982-2007. The location of the most predictive point from the reanalysis grid had to be assessed (closest points: 5\textdegree\~-~7.5\textdegree\ longitude and 45\textdegree\~-~47.5\textdegree\ latitude), as well an eventual time lapse between both series, due to the significant distance separating the weather stations and the reanalysis grid point. The most relevant point will probably not be the same for all stations, since we can expect to find the optimum in the direction of the main moisture inflows for the station. This assessment must be carried out for each station.

@@

The impact of the STW on the prediction performance has never been assessed before, due to the shortcoming of long precipitation series at a sub-daily time step. We could evaluate the impact of the STW approach on a 25 years time series with a high temporal resolution. After the introduction of the STW, the performance scores increased of the same order of magnitude as if we doubled the length of the archive, in this case study.

>>>>

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent to a finer resolution. Therefore, the analogues situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations do not occur systematically at the same time of the day, and that better candidates can be found at different times. On this assumption, a sliding time window (STW) was introduced to allow searching for candidates at different hours of the day.

@@

One can question the interest of using daily totals when a 6-hourly precipitation series can be used. The first reason is that the 6-hourly time series generated by the AM may not represent well the dynamic of the accurate precipitation (results not shown), due to a smoothed signal. Then, sometimes one just does not need finer resolution than the daily time step. Finally, when using a reconstructed precipitation archive, the errors in intra-day precipitation distributions have a lesser impact.

@@

It can be assumed that the atmospheric conditions with a low dynamism, such as the frequent anticyclonic situations, will not be radically improved by the introduction of the STW. Conversely, dynamic situations, such as weather disturbances, have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure area and the fronts change significantly during a day. We can therefore expect to improve particularly these situations with a higher dynamism when introducing a STW, as better matches to the target situation may be found.

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is. The results of this analysis are summarized in Figure \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation between two thresholds. The number of analogues being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations and confirms our intuition.

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the candidates situations are not more numerous, as they are subsampled in the previously $N_{1}$ selected analogues, but the dates may have changed. In contrast to earlier, both a reduction and an increase of the RMSE criterion values are possible. Figure \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

@@

A systematic improvement of the S1 values was previously observed. However, finding better analogue situations does not obligatory imply better forecast skills. Therefore, the impact of the introduction of the STW, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see section \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

The values of the CRPSS scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration. With the introduction of the STW, the performance loss related to the reduction of the archive length is retrieved. In our case, this improvement is approximately a doubling of the archive size.

@@

As demonstrated in the previous section, the performance improvement brought by the STW is lost when using a simplistic precipitation time series. In order to get closer to the chronology of the accurate precipitation, the reconstruction procedure must rely on some informative variable. Data from the NCEP/NCAR reanalysis 1 (section \ref{sec:data}) was used in this perspective, despite their rough resolution.

The first step consisted in determining which variables from the precipitable water or the relative humidity (at 1000~hPa, 925~hPa or 850~hPa) produces the precipitation chronology that is the most correlated (on non-zero values) with the accurate 6-hourly precipitation time series on the period 1982-2007. The location of the most predictive point from the reanalysis grid had to be assessed (closest points: 5\textdegree\~-~7.5\textdegree\ longitude and 45\textdegree\~-~47.5\textdegree\ latitude), as well an eventual time lapse between both series, due to the significant distance separating the weather stations and the reanalysis grid point. The most relevant point will probably not be the same for all stations, since one can expect to find the optimum in the direction of the main moisture inflows for the station. This assessment must be carried out for each station.

@@

The impact of the STW on the prediction performance has never been assessed before, due to the shortcoming of long precipitation series at a sub-daily time step. The impact of the STW approach could be assessed on a 25 years time series with a high temporal resolution. After the introduction of the STW, the performance scores increased of the same order of magnitude as if the length of the archive doubled, in this case study.
